---
output: 
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
geometry: margin=2.54cm
title: "Insert title of project here"
subtitle: "Web address for GitHub repository"
author: "Emily Wood"
fontsize: 12pt
mainfont: Times New Roman

---

\newpage
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

```{r setup, include=FALSE}
#This code chunk will tidy your knit PDF files, wrapping long code lines
#For it to work, the "formatR" package needs to be installed

#install.packages('formatR')

knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80), tidy=TRUE)
```

```{r setup, include=FALSE}
# Set your working directory

getwd()

# Load your packages

library('formatR')
library(tidyverse)
library(agricolae)
library(ggplot2)
library(lubridate)
library(rworldmap)


# Set your ggplot theme

mynewtheme <- theme_grey(base_size = 12) +
  theme(axis.text = element_text(color = "Dark Green"), 
        legend.position = "top")

theme_set(mynewtheme)

# Load your datasets

CITESPoaching2016 <- read_csv("/home/guest/EDA_2022/EDA-Fall2022/Project/comptab_2018-01-29 16_00_comma_separated.csv")
```


```{r setup, include=FALSE}

colnames(CITESPoaching2016)

```

```{r}
#datacleaning 

#rename columns 

colnames(CITESPoaching2016)[which(names(CITESPoaching2016) == "Importer reported quantity")] <- "Importer_reported_quantity"

colnames(CITESPoaching2016)[which(names(CITESPoaching2016) == "Exporter reported quantity")] <- "Exporter_reported_quantity"

colnames(CITESPoaching2016)[which(names(CITESPoaching2016) == "App.")] <- "App"

#fix NAs 

#CITESPoaching2016["Importer_reported_quantity"][is.na(CITESPoaching2016["Importer_reported_quantity"])] <- 0

#CITESPoaching2016["Exporter_reported_quantity"][is.na(CITESPoaching2016["Exporter_reported_quantity"])] <- 0

#create one reported quantity column 

CITESPoaching2016$Reported_quantity <- CITESPoaching2016$Importer_reported_quantity + CITESPoaching2016$Exporter_reported_quantity 

```


```{r}

c1<- ggplot(CITESPoaching2016, aes(x=Exporter_reported_quantity ))+ geom_histogram(binwidth=50, colour="blue", fill="light blue", alpha=0.3)+ xlim(0,1500)+ylim(0,2500)+
  labs(title="Export Distribution") 
c1

c2<- ggplot(CITESPoaching2016, aes(x=Importer_reported_quantity ))+ geom_histogram(binwidth=50, colour="dark green", fill="light green", alpha=0.3)+ xlim(0,1500)+ylim(0,2500)+
  labs(title="Import Distribution") 
c2

CITESPoaching2016["Importer_reported_quantity"][is.na(CITESPoaching2016["Importer_reported_quantity"])] <- 0

CITESPoaching2016["Exporter_reported_quantity"][is.na(CITESPoaching2016["Exporter_reported_quantity"])] <- 0


Graphnew <- 
  ggplot(CITESPoaching2016, aes(x= Exporter_reported_quantity, y= Importer_reported_quantity,color = App, alpha = .3))+
  geom_point() +
  xlim(0, 4000000) +
  ylim(0, 4000000)

print(Graphnew)   


Graph1test <- 
  ggplot(subset(CITESPoaching2016, App == "I"), aes(x= Exporter_reported_quantity, y= Importer_reported_quantity,color = App, alpha = .3)) +
  geom_point() 
print(Graph1test)

Graphnew2 <- 
  ggplot(CITESPoaching2016, aes(x= Importer_reported_quantity, y= Exporter_reported_quantity ))+
  geom_point(aes(color = App, alpha = .3)) +
  xlim(0, 4000000) +
  ylim(0, 4000000)

print(Graphnew2) 

Graphnew3 <- 
  ggplot(CITESPoaching2016, aes(Reported_quantity, color = Class))+
  geom_point() +
  xlim(0, 4000000) +
  ylim(0, 4000000)

print(Graphnew3) 

test <- ggplot(CITESPoaching2016, aes(x= Importer_reported_quantity, y= Exporter_reported_quantity)) + geom_point()

test



```


# Rationale and Research Questions

Does level of protection, class, term affect the amount exported/imported?

What might the best set of predictors for Poaching?


```{r}

#Two way ANOVA

Poaching.2way2 <- lm(data = CITESPoaching2016, Reported_quantity ~ Class + App)
summary(Poaching.2way2)

Testregression <- lm(data = CITESPoaching2016, Reported_quantity ~ Class)

summary(Testregression)

```
```{r}

CITESAIC <- lm(data = CITESPoaching2016, Importer_reported_quantity ~ Class + App + Taxon + Term)

step(CITESAIC)

```


\newpage

# Dataset Information



\newpage

# Exploratory Analysis 



\newpage

# Analysis



## Question 1: <insert specific question here and add additional subsections for additional questions below, if needed>

## Question 2: 




\newpage

# Summary and Conclusions


\newpage

# References
<add references here if relevant, otherwise delete this section> 
