---
output: 
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
geometry: margin=2.54cm
title: "Insert title of project here"
subtitle: "Web address for GitHub repository"
author: "Name"
fontsize: 12pt
mainfont: Times New Roman

---

\newpage
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
install.packages("trend")
library(trend)
install.packages("zoo")
library(zoo)
install.packages("Kendall")
library(Kendall)
install.packages("tseries")
library(tseries)
install.packages("readxl")
library("readxl")

# Set theme
mytheme <- theme_classic(base_size = 14) +
  theme(axis.text = element_text(color = "black"), 
        legend.position = "top")
theme_set(mytheme)

#read in data 

NP_Visitation <- read_excel("/home/guest/EDA_2022/EDA-Fall2022/Project/Annual Visitation By Park (1979 - Last Calendar Year) (1).xlsx")

```


# Rationale and Research Questions

How does NP visitation change over time 

Is national park visitation affected by seasonality 


```{r}

```


\newpage

# Dataset Information



\newpage

# Exploratory Analysis 



\newpage

# Analysis



## Question 1: <insert specific question here and add additional subsections for additional questions below, if needed>

## Question 2: 




\newpage

# Summary and Conclusions


\newpage

# References
<add references here if relevant, otherwise delete this section> 
