---
title: "Assignment 2: Coding Basics"
author: "Emily Wood"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(tidy.opts = list(width.cutoff=80), tidy = TRUE)

```


## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on coding basics.

## Directions

1.  Rename this file `<FirstLast>_A02_CodingBasics.Rmd` (replacing `<FirstLast>` with your first and last name).
2.  Change "Student Name" on line 3 (above) with your name.
3.  Work through the steps, **creating code and output** that fulfill each instruction.
4.  Be sure to **answer the questions** in this assignment document.
5.  When you have completed the assignment, **Knit** the text and code into a single PDF file.
6.  After Knitting, submit the completed exercise (PDF file) to Sakai.

## Basics Day 1

1.  Generate a sequence of numbers from one to 100, increasing by fours. Assign this sequence a name.

2.  Compute the mean and median of this sequence.

3.  Ask R to determine whether the mean is greater than the median.

4.  Insert comments in your code to describe what you are doing.

```{r}
#1.
 seq1 <- seq(1,100,4)

#Seq1 is a sequence form 1 to 100 that increases by four

#2. 

meanseq1 <- mean(seq1)
meanseq1

medianseq1 <- median(seq1) 
medianseq1

#Here I am calculating the mean and median of seq1 and assigning them a name. Both the mean and median are 49.

#3. 

meanseq1 > medianseq1

#Here I am asking R to determin if the mean is greater than the mdeian. It is not so the result is 'False'. 

```

## Basics Day 2

5.  Create a series of vectors, each with four components, consisting of (a) names of students, (b) test scores out of a total 100 points, and (c) whether or not they have passed the test (TRUE or FALSE) with a passing grade of 50.

6.  Label each vector with a comment on what type of vector it is.

7.  Combine each of the vectors into a data frame. Assign the data frame an informative name.

8.  Label the columns of your data frame with informative titles.

```{r}

#5.

names_of_students <- c("Kelly", "Emily", "Sam", "Meech")
class(names_of_students)
names_of_students

#This is the vector of the names of each student. It is a character vector.

test_scores <- c(11, 22, 77, 88)
class(test_scores)
test_scores

#This is the vector of the test scores of each student. It is a numeric vector.

Pass_Fail <- c(FALSE, FALSE, TRUE, TRUE)
class(Pass_Fail)
Pass_Fail

#This is the vector of whether each student passed or not. It is a logical vector.

#7. 

df_Pass_Fail <- as.data.frame(Pass_Fail)

dataFrame_TestResults <- cbind(df_Pass_Fail,test_scores,names_of_students)
dataFrame_TestResults
is.data.frame(dataFrame_TestResults)

#8.

dataFrame_TestResults_Final<- data.frame("Name"=names_of_students,"Grade"=test_scores,"Passed" =Pass_Fail)
dataFrame_TestResults_Final

```

9.  QUESTION: How is this data frame different from a matrix?

> Answer: Matrices only contain one type of data. Dataframes like this one can contain multiple types such as numeric, character and logical verticies.

10. Create a function with an if/else statement. Your function should take a **vector** of test scores and print (not return) whether a given test score is a passing grade of 50 or above (TRUE or FALSE). You will need to choose either the `if` and `else` statements or the `ifelse` statement. 

11. Apply your function to the vector with test scores that you created in number 5.

```{r}

gradingrecipe <- function(x){
  ifelse(x>50, TRUE, FALSE) 
  
}

Gradingmeal <-gradingrecipe(test_scores)

Gradingmeal

```

12. QUESTION: Which option of `if` and `else` vs. `ifelse` worked? Why?

> Answer: 'ifelse' worked because it checks every element of a vector. The 'if' and 'else' staments only deal with a single value. Because of this, R showed an error when I tried using a function with 'if' and 'else' statements with the test scores vector. 
