---
title: "Assignment 09: Data Scraping"
author: "Emily Wood"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on data scraping. 

## Directions
1. Rename this file `<FirstLast>_A09_DataScraping.Rmd` (replacing `<FirstLast>` with your first and last name).
2. Change "Student Name" on line 3 (above) with your name.
3. Work through the steps, **creating code and output** that fulfill each instruction.
4. Be sure to **answer the questions** in this assignment document.
5. When you have completed the assignment, **Knit** the text and code into a single PDF file.

```{r setup, include=FALSE}

knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80), tidy=TRUE)
```

## Set up 
1. Set up your session:

* Check your working directory
* Load the packages `tidyverse`, `rvest`, and any others you end up using.
* Set your ggplot theme

```{r, message = FALSE}
#1

getwd()
library(tidyverse)
library(lubridate)
library(viridis)

install.packages("rvest")
library(rvest)

install.packages("dataRetrieval")
library(dataRetrieval)

install.packages("tidycensus")
library(tidycensus)

# Set theme
mytheme <- theme_classic() +
  theme(axis.text = element_text(color = "black"), 
        legend.position = "top")
theme_set(mytheme)

```

2. We will be scraping data from the NC DEQs Local Water Supply Planning website, specifically the Durham's 2021 Municipal Local Water Supply Plan (LWSP): 
 * Navigate to https://www.ncwater.org/WUDC/app/LWSP/search.php
 * Scroll down and select the LWSP link next to Durham Municipality. 
 * Note the web address: <https://www.ncwater.org/WUDC/app/LWSP/report.php?pwsid=03-32-010&year=2021>
 
Indicate this website as the as the URL to be scraped. (In other words, read the contents into an `rvest` webpage object.)

```{r set.the.scraping.website}
#2

webpage <- read_html("https://www.ncwater.org/WUDC/app/LWSP/report.php?pwsid=03-32-010&year=2021")
webpage

```

3. The data we want to collect are listed below:

* From the "1. System Information" section:
 * Water system name
 * PSWID
 * Ownership
 
* From the "3. Water Supply Sources" section:
 * Maximum Daily Use (MGD) - for each month

In the code chunk below scrape these values, assigning them to four separate variables.

>HINT: The first value should be "Durham", the second "03-32-010", the third "Municipality", and the last should be a vector of 12 numeric values (represented as strings), with the first value being "27.6400".

```{r scrape.the.data}
#3
water.system.name <- webpage %>%
html_nodes("div+ table tr:nth-child(1) td:nth-child(2)") %>%
html_text()
  
pwsid <- webpage %>%
html_nodes("td tr:nth-child(1) td:nth-child(5)") %>%
html_text()
  
ownership <- webpage %>%
html_nodes("div+ table tr:nth-child(2) td:nth-child(4)") %>%
html_text()
  
max.withdrawals.mgd <- webpage %>%
html_nodes("th~ td+ td") %>%
html_text()

```


4. Convert your scraped data into a dataframe. This dataframe should have a column for each of the 4 variables scraped and a row for the month corresponding to the withdrawal data. Also add a Date column that includes your month and year in data format. (Feel free to add a Year column too, if you wish.)

>TIP: Use `rep()` to repeat a value when creating a dataframe.

>NOTE: It's likely you won't be able to scrape the monthly widthrawal data in chronological order. You can overcome this by creating a month column manually assigning values in the order the data are scraped: "Jan", "May", "Sept", "Feb", etc...

5. Create a line plot of the maximum daily withdrawals across the months for 2021

```{r create.a.dataframe.from.scraped.data}
#4

Month <- c(1, 5, 9, 2, 6, 10, 3, 7, 11, 4, 8, 12)
Year <- c(2021)

the_df <- data.frame(WaterSystem = water.system.name, PWSID = pwsid, Ownership = ownership,
`Max Day Use` = as.numeric(max.withdrawals.mgd)) %>%
mutate(Month = Month, Year = Year, Date = my(paste0(Month, "-", Year)), PWSID = !!pwsid,
Ownership = !!ownership)

the_df

#5
plot1 <- ggplot(the_df, aes(x = Date, y = Max.Day.Use)) + 
    geom_line(aes(group = 1)) + 
    geom_smooth(method = "loess", se = FALSE, aes(color = "salmon")) + 
    labs(title = paste("Maximum Daily Withdrawals per Months for 2021"),
    subtitle = "Durham", y = "Max Day Use (MGD)", x = "Date")

plot1

```

6. Note that the PWSID and the year appear in the web address for the page we scraped. Construct a function using your code above that can scrape data for any PWSID and year for which the NC DEQ has data. **Be sure to modify the code to reflect the year and site (pwsid) scraped**.

```{r construct.a.scraping.function}
#6.

base_url <- "https://www.ncwater.org/WUDC/app/LWSP/report.php?pwsid="
pwsid <- "03-32-010"
Year <- 2015
scrape_url <- paste0(base_url, pwsid, "&year=", Year)
website <- read_html(scrape_url)


scrape.it <- function(pwsid, Year) {
  
website <- read_html(paste0("https://www.ncwater.org/WUDC/app/LWSP/report.php?pwsid=",pwsid, "&year=", Year)) 

water.system.name <- website %>% html_nodes("div+ table tr:nth-child(1) td:nth-child(2)") %>% html_text()

pwsid <- website %>% html_nodes("td tr:nth-child(1) td:nth-child(5)") %>% html_text()

ownership <- website %>% html_nodes("div+ table tr:nth-child(2) td:nth-child(4)") %>% html_text()

max.withdrawals.mgd <- website %>% html_nodes("th~ td+ td") %>% html_text()

the_df2 <- data.frame(WaterSystem = water.system.name, 
                      PWSID = pwsid, Ownership = ownership, 
                      `Max Day Use` = as.numeric(max.withdrawals.mgd)) %>% 
 mutate(Month = Month, 
         Year = Year, 
         Date = my(paste0(Month, "-", Year)), 
         PWSID = !!pwsid,
        Ownership = !!ownership)

  return(the_df2)
}

```

7. Use the function above to extract and plot max daily withdrawals for Durham (PWSID='03-32-010') for each month in 2015

```{r fetch.and.plot.Durham.2015.data}
#7

Durham2015_df <- scrape.it("03-32-010", 2015)
print(Durham2015_df)

plot2 <- ggplot(Durham2015_df, aes(x = Date, y = Max.Day.Use)) + 
    geom_line(aes(group = 1)) + 
    geom_smooth(method = "loess", se = FALSE, aes(color = "salmon")) + 
    labs(title = paste("Maximum Daily Withdrawals per Months for", Year),
    subtitle = "Durham", y = "Max Day Use (MGD)", x = "Date")

plot2

```

8. Use the function above to extract data for Asheville (PWSID = 01-11-010) in 2015. Combine this data with the Durham data collected above and create a plot that compares Asheville's to Durham's water withdrawals.

```{r fetch.and.plot.Asheville.2015.data}
#8

Ashville2015 <- scrape.it("01-11-010", 2015)
print(Ashville2015)

AshvilleDurham<- rbind(Durham2015_df, Ashville2015)
print(AshvilleDurham)

plot3 <- ggplot(AshvilleDurham) + 
    geom_line(aes(x = Date, y = Max.Day.Use, color = WaterSystem)) +
    labs(title = paste("Durham and Asheville Monthly Maximum Daily Withdrawals in 2015"),
    subtitle = "Emily Wood", y = "Max Day Use", x = "Date")

plot3
```


9. Use the code & function you created above to plot Asheville's max daily withdrawal by months for the years 2010 thru 2019.Add a smoothed line to the plot.

>TIP: See Section 3.2 in the "09_Data_Scraping.Rmd" where we apply "map2()" to iteratively run a function over two inputs. Pipe the output of the map2() function to `bindrows()` to combine the dataframes into a single one. 

```{r fetch.and.plot.Asheville.multiyear.data}
#9

years = rep(2010:2019)
pwsidnew = "01-11-010"
df_10_19 <- years %>%
  map(scrape.it, pwsid = pwsidnew) %>%
  bind_rows()

plot4 <- ggplot(df_10_19, aes(y = Max.Day.Use, x = Date)) + 
  geom_line() + 
  geom_smooth(method = "loess",se = FALSE, aes(color = "salmon")) + 
  labs(title = paste("Monthly Max Daily Withdrawals from 2010 tp 2019 "),
  subtitle = "Asheville", y = "Max Day Use", x = "Date")

plot4

```

>Question: Just by looking at the plot (i.e. not running statistics), does Asheville have a trend in water usage over time?

Just from looking at the plot I believe that Asheville has rising (positive) trend in max daily water usage overtime. This is depicted in the upward curve of the trend line. 