---
title: "Assignment 3: Data Exploration"
author: "Emily Wood"
output: pdf_document
geometry: margin=2.54cm
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on Data Exploration.

## Directions

1.  Rename this file `<FirstLast>_A03_DataExploration.Rmd` (replacing `<FirstLast>` with your first and last name).
2.  Change "Student Name" on line 3 (above) with your name.
3.  Work through the steps, **creating code and output** that fulfill each instruction.
4.  Be sure to **answer the questions** in this assignment document.
5.  When you have completed the assignment, **Knit** the text and code into a single PDF file.
6.  After Knitting, submit the completed exercise (PDF file) to the dropbox in Sakai.

The completed exercise is due on Sept 30th.

## Set up your R session

1.  Check your working directory, load necessary packages (tidyverse), and upload two datasets: the ECOTOX neonicotinoid dataset (ECOTOX_Neonicotinoids_Insects_raw.csv) and the Niwot Ridge NEON dataset for litter and woody debris (NEON_NIWO_Litter_massdata_2018-08_raw.csv). Name these datasets "Neonics" and "Litter", respectively. Be sure to include the subcommand to read strings in as factors.

```{r,  message = FALSE}
getwd()
#install.packages("tidyverse")
Neonics <- read.csv("./Data/Raw/ECOTOX_Neonicotinoids_Insects_raw.csv", stringsAsFactors = TRUE)
Litter <- read.csv("./Data/Raw/NEON_NIWO_Litter_massdata_2018-08_raw.csv", stringsAsFactors = TRUE)

knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80), tidy=TRUE)
```

## Learn about your system

2.  The neonicotinoid dataset was collected from the Environmental Protection Agency's ECOTOX Knowledgebase, a database for ecotoxicology research. Neonicotinoids are a class of insecticides used widely in agriculture. The dataset that has been pulled includes all studies published on insects. Why might we be interested in the ecotoxicology of neonicotinoids on insects? Feel free to do a brief internet search if you feel you need more background information.

> Answer: We would be interested in the ecotoxicology of neonicotinoids on insects to see the results of the pesticides. Were the pesticides effective, were there any unforeseen consequences such as behavioral changes or emergence patterns. It is also important to know how these chemicals affect insects that were not the target pest. 

3.  The Niwot Ridge litter and woody debris dataset was collected from the National Ecological Observatory Network, which collectively includes 81 aquatic and terrestrial sites across 20 ecoclimatic domains. 32 of these sites sample forest litter and woody debris, and we will focus on the Niwot Ridge long-term ecological research (LTER) station in Colorado. Why might we be interested in studying litter and woody debris that falls to the ground in forests? Feel free to do a brief internet search if you feel you need more background information.

> Answer: Studying litter and woody debris would give insight to how the organic material of soil is formed over time. It also shows the composition for habitat for numerous species including insects, invertebrates and other smaller organisms. Litter and debris also act as a barrier between soil and the elements. It prevents the soil itself from drying out or eroding away during precipitation events. 

4.  How is litter and woody debris sampled as part of the NEON network? Read the NEON_Litterfall_UserGuide.pdf document to learn more. List three pieces of salient information about the sampling methods here:

> Answer: 
 1. Litter and fine woody debris are collected from elevated traps and ground traps.
 2.All masses measured are reported at the spatial resolution of a single trap and the temporal resolution of a single collection event. 
 3. Mass data for each collection event are measured separately for different groups. Examples of these groups include leaves, twigs, Needles, Seeds, etc. 


## Obtain basic summaries of your data (Neonics)

5.  What are the dimensions of the dataset?

```{r}

length(Neonics)
dim(Neonics)
colnames(Neonics)
str(Neonics)

#There are 30 columns and 4,623 entries in the Neonics dataframe.

```

6.  Using the `summary` function on the "Effect" column, determine the most common effects that are studied. Why might these effects specifically be of interest?

```{r}
summary(Neonics$Effect)


```

> Answer: The most commonly studied effects are mortality and population. It makes sense that mortality is studied as that is often the desired outcome of pesticides. Population also makes sense especially if a pest is overpopulated in an area. By studying population you can assess how it fluctuates based on the use of this pesticide in an area. 

7.  Using the `summary` function, determine the six most commonly studied species in the dataset (common name). What do these species have in common, and why might they be of interest over other insects? Feel free to do a brief internet search for more information if needed.

```{r}
summary(Neonics$Species.Common.Name)
```

> Answer: The most commonly studied species is the Honey Bee and two types of bumble bees. They are likely studied because they are important crop and native plant pollinators. If these pesticides are killing them at a high enough rate this could cause a collapse in our food system. 

8.  Concentrations are always a numeric value. What is the class of Conc.1..Author. in the dataset, and why is it not numeric?

```{r}
class(Neonics$Conc.1..Author.)

```

> Answer: It is not currently numeric because we set everything as a factor when we read in the csv file. 

## Explore your data graphically (Neonics)

9.  Using `geom_freqpoly`, generate a plot of the number of studies conducted by publication year.

```{r}
library("ggplot2")

ggplot(Neonics) +
  geom_freqpoly(aes(Publication.Year), bins = 20)

```

10. Reproduce the same graph but now add a color aesthetic so that different Test.Location are displayed as different colors.

```{r}

ggplot(Neonics) +
  geom_freqpoly(aes(x = Publication.Year, color = Test.Location), bins = 20) 
  
```

Interpret this graph. What are the most common test locations, and do they differ over time?

> Answer: The most common test locations are "Field Natural" and "Lab". According to the graph the "Field Natural" test location was slightly more common through the 90s. In the early 2000s the Lab became the most common test location with "field Natural" spiking one more time in 2010. 

11. Create a bar graph of Endpoint counts. What are the two most common end points, and how are they defined? Consult the ECOTOX_CodeAppendix for more information.

```{r fig.height = 3, fig.width = 4}

ggplot(Neonics, aes(x = Endpoint)) +
  geom_bar()

summary(Neonics$Endpoint)

```

> Answer: The two most popular endpoints are NOEL and LOEL. NOEL stands for no observable effect level. The highest concentration does not show effects that differ from control groups. LOEL stands for lowest observable effect level. This means that the lowest dose
producing effects were significantly different from the controls. 

## Explore your data (Litter)

12. Determine the class of collectDate. Is it a date? If not, change to a date and confirm the new class of the variable. Using the `unique` function, determine which dates litter was sampled in August 2018.

```{r}

class(Litter$collectDate)
Litter$collectDate <- as.Date(Litter$collectDate, format = "%Y-%m-%d") 

unique(Litter$collectDate, 2018-08)

#Litter was samples twice in AUGUST OF 2018. The collection dates were on the 2nd and 30th.
  
```

13. Using the `unique` function, determine how many plots were sampled at Niwot Ridge. How is the information obtained from `unique` different from that obtained from `summary`?

```{r}

unique(Litter$plotID)

```

> Answer: 12 plots were sampled at Niwot Ridge. The unique function gives us a count of how many variables are in the vector. The summary function tells us how many times each variable is listed in the vector. 

14. Create a bar graph of functionalGroup counts. This shows you what type of litter is collected at the Niwot Ridge sites. Notice that litter types are fairly equally distributed across the Niwot Ridge sites.

```{r}

ggplot(Litter, aes(x = functionalGroup)) +
  geom_bar()

```

15. Using `geom_boxplot` and `geom_violin`, create a boxplot and a violin plot of dryMass by functionalGroup.

```{r}

ggplot(Litter) +
  geom_boxplot(aes(x = dryMass, y = functionalGroup))

ggplot(Litter) +
  geom_violin(aes(x = dryMass, y = functionalGroup))

```

Why is the boxplot a more effective visualization option than the violin plot in this case?

> Answer: Boxplots show summary statistics. Violin plots take that one step further and also show density. In this case, the density does not compound enough to show any trends in dryMass. We can see some areas of higher density at 0 for seeds and leaves but there is not enough information to give us the recognizable box plot waves. 

What type(s) of litter tend to have the highest biomass at these sites?

> Answer:According to the boxplots, needles tend to have the highest biomass at these sites. This is followed by the mixed category and twigs category. 

