---
title: "Assignment 6: GLMs (Linear Regressios, ANOVA, & t-tests)"
author: "Student Name"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
#This code chunk will tidy your knit PDF files, wrapping long code lines
#For it to work, the "formatR" package needs to be installed

#install.packages('formatR')

knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80), tidy=TRUE)
```

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on generalized linear models. 

## Directions
1. Rename this file `<FirstLast>_A06_GLMs.Rmd` (replacing `<FirstLast>` with your first and last name).
2. Change "Student Name" on line 3 (above) with your name.
3. Work through the steps, **creating code and output** that fulfill each instruction.
4. Be sure to **answer the questions** in this assignment document.
5. When you have completed the assignment, **Knit** the text and code into a single PDF file.


## Set up your session 
1. Set up your session. Check your working directory. Load the tidyverse, agricolae and other needed packages. Import the *raw* NTL-LTER raw data file for chemistry/physics (`NTL-LTER_Lake_ChemistryPhysics_Raw.csv`). Set date columns to date objects.

2. Build a ggplot theme and set it as your default theme.

```{r setup2}
#1

getwd()
library('formatR')
library(tidyverse)
library(agricolae)
library(ggplot2)
library(lubridate)

ChemPhys_Raw <- 
  read.csv("./Data/Raw/NTL-LTER_Lake_ChemistryPhysics_Raw.csv",stringsAsFactors = T)


ChemPhys_Raw$sampledate <- as.Date(ChemPhys_Raw$sampledate, format = "%m/%d/%y")
class(ChemPhys_Raw$sampledate)

#2

mynewtheme <- theme_grey(base_size = 12) +
  theme(axis.text = element_text(color = "Dark Green"), 
        legend.position = "top")

theme_set(mynewtheme)

```

## Simple regression
Our first research question is: Does mean lake temperature recorded during July change with depth across all lakes?

3. State the null and alternative hypotheses for this question:
> Answer:
H0: Mean lake temperature recorded during July changes with depths across all lakes.
Ha: Mean lake temperature recorded during July does not change with depths across all lakes.


4.  Wrangle your NTL-LTER dataset with a pipe function so that the records meet the following criteria: 
 * Only dates in July. 
 * Only the columns: `lakename`, `year4`, `daynum`, `depth`, `temperature_C`
 * Only complete cases (i.e., remove NAs)

5. Visualize the relationship among the two continuous variables with a scatter plot of temperature by depth. Add a smoothed line showing the linear model, and limit temperature values from 0 to 35 °C. Make this plot look pretty and easy to read.

```{r scatterplot}
#4

NTL_New <- 
  ChemPhys_Raw %>%
  mutate(month = month(sampledate), year = year(sampledate)) %>%
  filter(month == "7") %>%
  select( lakename, year4, daynum, depth, temperature_C )%>%
  drop_na()

#5

Graphnew <- 
  ggplot(NTL_New, aes (x= depth, y= temperature_C))+
  geom_point(aes(color = temperature_C), alpha =0.3) +
  geom_smooth(method=lm, color= "black" )+
  xlim(0, 20) +
  ylim(0, 35)+
  xlab("Depth in Meters")+
  ylab("Temperature Celsius")
print(Graphnew)

```


6. Interpret the figure. What does it suggest with regards to the response of temperature to depth? Do the distribution of points suggest about anything about the linearity of this trend?

> Answer: According to the graph as depth increases temperature decreases. The distribution suggests a negatively correlated relationship. We can see a linear trend in the data but it does appear to level off past 6 meters and is also less linear at the least shallow depths. The relationship appears to be most linear from around 2 to 7 meters.


7. Perform a linear regression to test the relationship and display the results

```{r linear.regression}
#7

NTL.regression <- lm(data = NTL_New, temperature_C ~ depth)

summary(NTL.regression)

cor.test(NTL_New$temperature_C, NTL_New$depth)

```


8. Interpret your model results in words. Include how much of the variability in temperature is explained by changes in depth, the degrees of freedom on which this finding is based, and the statistical significance of the result. Also mention how much temperature is predicted to change for every 1m change in depth. 

> Answer: The p value is very small so that means it is statistically significant. This means that we accept our alternative hypothesis that mean lake temperature recorded during July does change with depths across all lakes. The degree of freedom is 9726. Temperature will change -1.95 degrees for every unit of increase in depth. 73.9% of the variability of temperature is explained by depth. 


---

## Multiple regression
Let's tackle a similar question from a different approach. Here, we want to explore what might the best set of predictors for lake temperature in July across the monitoring period at the North Temperate Lakes LTER. 


9. Run an AIC to determine what set of explanatory variables (year4, daynum, depth) is best suited to predict temperature.

10. Run a multiple regression on the recommended set of variables. 

```{r temperature.model}
#9


NLT_NewAIC <- lm(data = NTL_New, temperature_C ~ depth + year4 + daynum)
step(NLT_NewAIC)

#10

NLT_MR <- lm(data = NTL_New, temperature_C ~ depth + daynum)

summary(NLT_MR)

par(mfrow = c(2,2), mar=c(4,4,4,4))
plot(NLT_MR)
par(mfrow = c(1,1))

```

11. What is the final set of explanatory variables that the AIC method suggests we use to predict temperature in our multiple regression? How much of the observed variance does this model explain? Is this an improvement over the model using only depth as the explanatory variable?

> Answer: The explanatory variables that the AIC method suggests we use to predict temperature are depth and day number. This model explains 74.1% of the variance in temperature. This is a slight improvement over the model using just depth which explained 73.9% of the variance. 

---
## Analysis of Variance

12. Now we want to see whether the different lakes have, on average, different temperatures in the month of July. Run an ANOVA test to complete this analysis. (No need to test assumptions of normality or similar variances.) Create two sets of models: one expressed as an ANOVA models and another expressed as a linear model (as done in our lessons).

```{r anova.model}
#12

LakeTemp.July.anova <- aov(data = NTL_New, temperature_C ~ lakename)
summary(LakeTemp.July.anova)

LakeTemp.July.anova2 <- lm(data = NTL_New, temperature_C ~ lakename)
summary(LakeTemp.July.anova2)

```

13. Is there a significant difference in mean temperature among the lakes? Report your findings. 

> Answer: Our p value for the AOC model is less than .05 so we reject our null hypothesis that all lakes have the same mean temperature. In other words the anoval model found that the difference between the pair and group means are statistically significant so we reject it. 


14. Create a graph that depicts temperature by depth, with a separate color for each lake. Add a geom_smooth (method = "lm", se = FALSE) for each lake. Make your points 50 % transparent. Adjust your y axis limits to go from 0 to 35 degrees. Clean up your graph to make it pretty. 

```{r scatterplot.2}
#14.

Graphnew2 <- 
  ggplot(NTL_New, aes (x= depth, y = temperature_C))+
  geom_point(aes(color = lakename), alpha = 0.5) +
  geom_smooth(method= "lm", se = FALSE, color = "blue")+
  xlim(0, 20) +
  ylim(0, 35)+
  xlab("Depth in Meters")+
  ylab("Temperature Celsius")
print(Graphnew2)

```

15. Use the Tukey's HSD test to determine which lakes have different means.

```{r tukey.test}
#15

TukeyHSD(LakeTemp.July.anova)

```

16.From the findings above, which lakes have the same mean temperature, statistically speaking, as Peter Lake? Does any lake have a mean temperature that is statistically distinct from all the other lakes?

>Answer: From the information provided in the tukey test Paul lake and Ward lake statstically dont have different mean temperatures from Peter lake. The p values given for the comparisons on the tukey test are all greater than .05 meaning the means aren't statistically different. Central long lake appears to have the most statistically distink mean average temperature. It hold a p-value of 0.0 for all comparisons. 


17. If we were just looking at Peter Lake and Paul Lake. What's another test we might explore to see whether they have distinct mean temperatures? 

>Answer: We could use the bartlett test to compare just Peter Lake and Paul Lake. This test check to see if the variances of the samples are the same. 


18. Wrangle the July data to include only records for Crampton Lake and Ward Lake. Run the two-sample T-test on these data to determine whether their July temperature are same or different. What does the test say? Are the mean temperatures for the lakes equal? Does that match you answer for part 16?

```{r t.test}

NTL_New_2 <- 
  NTL_New %>%
  filter(lakename == "Crampton Lake" | lakename == "Ward Lake")

ttestNTL <- t.test(data = NTL_New_2, temperature_C ~ lakename)
ttestNTL

```

>Answer: The p-value is greater than 0.5 (not significant) so that means we reject our null hypothesis that the mean temperature for both lakes is equal.The means that were reported in the t test are the same as the means you would get based off of the differences between means reported in the Tukey test. Meaning both tests gave us the same means. 
